---
output:
  word_document: default
  html_document: default
---
### Erica MacBride
## Dr. Hill, BAN 502
## Classification Trees

###Installing Packages
```{r}
#install.packages("rpart.plot")
#install.packages("rattle")
#install.packages("RColorBrewer")
```

### Library Packages
```{r}
library(tidyverse)
library(tidymodels)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(RColorBrewer)
```
### Loading and Cleaning Data
```{r}
parole <- read_csv("parole (1).csv")

parole = parole %>% mutate(male = as_factor(male)) %>% 
  mutate(male = fct_recode(male, "Female" = "0", "Male" = "1" ))

parole = parole %>% mutate(race = as_factor(race)) %>% 
  mutate(race = fct_recode(race, "Other" = "2", "White" = "1" ))

parole = parole %>% mutate(state = as_factor(state)) %>% 
  mutate(state = fct_recode(state, "Kentucky" = "2", "Louisiana" = "3", "Virginia" = "4", "Other"= "1" ))

parole = parole %>% mutate(crime = as_factor(crime)) %>% 
  mutate(crime = fct_recode(crime, "larceny" = "2", "drug-related crime" = "3", "driving-related crime"="4", "other crime" = "1" ))

parole = parole %>% mutate(multiple.offenses = as_factor(multiple.offenses)) %>% 
  mutate(multiple.offenses = fct_recode(multiple.offenses, "incarcerated-multiple" = "1", "otherwise" = "0" ))

parole = parole %>% mutate(violator = as_factor(violator)) %>% 
  mutate(violator = fct_recode(violator, "Violated" = "1", "Completed w no Violation" = "0" ))

```


### Task 1
```{r}
set.seed(12345) 
parole_split = initial_split(parole, prob = 0.70, strata = violator)
train = training(parole_split)
test = testing(parole_split)

```

### Task 2
```{r}
parole_recipe = recipe(violator ~ ., train)

tree_model = decision_tree() %>% 
  set_engine("rpart", model = TRUE) %>% 
  set_mode("classification")

parole_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(parole_recipe)

parole_fit = fit(parole_wflow, train)
```

### Task 3
```{r}
tree = parole_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")
```

```{r}
fancyRpartPlot(tree,tweak = 3)
```
The state is not other, kentucky or virginia, so no, Louisiana, yes to multiple offenses, less than 13 years, time served is less than 6.1 years, the parolee is less than 41 years so he falls into the completed with no violation.  


### Task 4
```{r}
parole_fit$fit$fit$fit$cptable
```
The optimal CP is .01, the tree from task 2 is associated with this optimal cp as it has 14 splits.  

### Task 5
```{r}
set.seed(123)
folds = vfold_cv(train, v = 5)
```
```{r}
parole_recipe = recipe(violator ~., train) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model = decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% #don't forget the model = TRUE flag
  set_mode("classification")

tree_grid = grid_regular(cost_complexity(),
                          levels = 25) #try 25 sensible values for cp

parole_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(parole_recipe)

tree_res = 
  parole_wflow %>% 
  tune_grid(
    resamples = folds,
    grid = tree_grid
    )

tree_res
```
```{r}
tree_res %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 
```
```{r}
best_tree = tree_res %>%
  select_best("accuracy")

best_tree
```

### Task 6
The optimal accuracy is a .1 value.  


### Task 7
```{r}
final_wf = 
  parole_wflow %>% 
  finalize_workflow(best_tree)
```

```{r}
final_fit = fit(final_wf, train)

tree = final_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

#fancyRpartPlot(tree, tweak = 1.5)
```
### Task 8

```{r}
448/507
```
88%



### Task 9
```{r}
blood <- read_csv("Blood.csv")
blood = blood %>% mutate(DonatedMarch = as_factor(DonatedMarch)) %>% 
  mutate(DonatedMarch = fct_recode(DonatedMarch, "No" = "0", "Yes" = "1" ))
#summary(blood)

```
```{r}
set.seed(1234) 
blood_split = initial_split(blood, prop = 0.70, strata = DonatedMarch)
train2 = training(blood_split)
test2 = testing(blood_split)

```

```{r}
set.seed(1234)
folds = vfold_cv(train2, v = 5)
```

```{r}
blood_recipe = recipe(DonatedMarch ~., train2) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model2 = decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% #don't forget the model = TRUE flag
  set_mode("classification")

tree_grid2 = grid_regular(cost_complexity(),
                          levels = 25) #try 25 sensible values for cp

blood_wflow = 
  workflow() %>% 
  add_model(tree_model2) %>% 
  add_recipe(blood_recipe)

tree_res2 = 
  blood_wflow %>% 
  tune_grid(
    resamples = folds,
    grid = tree_grid2
    )

tree_res2
```
```{r}
tree_res2 %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 
```
Somewhere between .015 to .019 looks like the most optimal cp value.
```{r}
best_tree2 = tree_res2 %>%
  select_best("accuracy")

best_tree2
```
 
### Task 10
```{r}
final_wf2 = 
  blood_wflow %>% 
  finalize_workflow(best_tree2)
```
```{r}
final_fit2 = fit(final_wf2, train2)

tree2 = final_fit2 %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

fancyRpartPlot(tree2, tweak = 1.5) 
```
### Task 11
Predicting on the training set
```{r}
treepred_train = predict(final_fit2, train2, type = "class")
head(treepred_train)
```


```{r}
confusionMatrix(treepred_train$.pred_class,train2$DonatedMarch,positive="Yes")
```


```{r}
treepred_test = predict(final_fit2, test2, type = "class")
head(treepred_test)
```

```{r}
confusionMatrix(treepred_test$.pred_class,test2$DonatedMarch,positive="Yes")
```
The training and testing both improved in accuracy. The improvement was not large, but overall performance is good.
