---
output:
  word_document: default
  html_document: default
---
## Erica MacBride
### BAN 502 Predictive
### Dr. Hill

#### Task 1
```{r}
#Tidyverse is already installed on my computer, no need to run install packages
library(tidyverse)
```
#### Task 2
```{r}
diamonddata = diamonds
#view(diamonddata)
```
There are 10 columns and 53,940 rows in this data set.  

#### Task 3
```{r}
ggplot(diamonddata, aes(x= carat, y=price)) + geom_point()
```
The bigger the carat size, the more expensive the diamond is.  

#### Task 4
```{r}
ggplot(diamonddata, aes(x= carat, y=price, color = cut)) + geom_point()
```
You can buy a fair cut diamond that is much bigger in carat size for the same price you could buy an ideal cut 2 carat diamond. The better the cut, the smaller the carat size due to high prices. The worse cut diamonds can be bought at a much bigger carat size and typically a lower price compared to nicer cuts.  

#### Task 5
```{r}
ggplot(diamonddata, aes(x= carat, y=price, color = cut)) + geom_point() + facet_wrap(~ color)
```
The nicer the color the more expensive the diamond. The better the cut, the more expensive the diamond. You can buy a larger carat thats a less nice color for cheaper than you could buy an ideal and very good colored diamond.  

#### Task 6
```{r}
inventory <- read_csv("InventoryData.csv")
#summary(inventory)
#str(inventory)
```
#### Task 7
```{r}
inventoryA <- inventory %>% filter(Supplier == "A")
```
3695 Rows are in this data frame.  

#### Task 8
```{r}
inventoryA = mutate(inventoryA, OnHandRatio = `On Hand` / `Annual Demand`)
```
This line of code creates a new column titled On Hand Ration and is formulated by dividing the On Hand variable over the Annual Demand variable.  

#### Task 9
```{r}
avg_cost <- inventory %>% group_by(`Item SKU`) %>% summarise(SKUAvgCost = mean(`Cost per Unit ($)`), .groups = 'drop')
```

#### Task 10
Given my experience in R, I have not found a particular topic too challenging. My main concern is remembering what line of code does what and troubleshooting when I get an error.  
