---
output:
  word_document: default
  html_document: default
---
### Erica MacBride
## Dr. Hill, BAN 502
## Random Forests

### Packages
```{r}
#install.packages("gridExtra")
#install.packages("vip")
#install.packages("ranger")
```



### Libraries 
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidymodels)
library(caret)
library(gridExtra) 
library(vip) 
library(ranger)
```
### Loading Data
```{r Naming Columns}
drug <- read_csv("drug_data-1.csv")
names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity",
"Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive",
"SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis",
"Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh",
"LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")
#str(drug)

```
```{r Changing Values}
drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"

```

```{r Factor Conversion}
drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44",
"45_54", "55_64", "65_"))) %>%
mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
mutate(Education = factor(Education, labels =
c("Under16", "At16", "At17", "At18", "SomeCollege",
"ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
mutate(Country = factor(Country,
labels = c("USA", "NewZealand", "Other", "Australia",
"Ireland","Canada","UK"))) %>%
mutate(Ethnicity = factor(Ethnicity,
labels = c("Black", "Asian", "White", "White/Black", "Other",
"White/Asian", "Black/Asian"))) %>%
mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
select(-ID)
#str(drug_clean)
```
```{r}
drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% select(!(Semer:VSA))
names(drug_clean)

```

### Task 1
```{r}
summary(drug_clean)
```
There is no missingness. 

### Task 2
```{r}
set.seed(1234) 
drug_split = initial_split(drug_clean, prop = 0.7, strata = Nicotine) #70% in training
train = training(drug_split)
test = testing(drug_split)
```

### Task 3
```{r}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill")
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill")
grid.arrange(p1,p2,p3,p4)
```
```{r}
p1 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Nicotine, y=Nscore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y=Escore)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y=Oscore)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4)
```
```{r}
p1 = ggplot(train, aes(x = Nicotine, y=Ascore)) + geom_boxplot()
p2 = ggplot(train, aes(x = Nicotine, y=Cscore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y=Impulsive)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y=SS)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4)
```


### Task 4
```{r}
set.seed(123)
rf_folds = vfold_cv(train, v=5)
```

```{r}
drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% #add tuning of mtry and min_n parameters
  #setting trees to 100 here should also speed things up a bit, but more trees might be better
  set_engine("ranger", importance = "permutation") %>% #added importance metric
  set_mode("classification")

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)

rf_grid = grid_regular(
  mtry(range = c(2,8)),
  min_n(range = c(2,20)),
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = rf_grid 
)
```
```{r}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```

### Task 5
```{r}
best_rf = select_best(rf_res_tuned, "accuracy")

final_rf = finalize_workflow(
  drug_wflow,
  best_rf
)

final_rf
```
```{r}
final_rf_fit = fit(final_rf, train)
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```
The most important variable is Sensation Seeking (SS). The country is also important, specifically, the UK. Oscore is also an important variable. This makes sense to me as oscore is represenative of openess to experience which in the real world I feel as if this ties well to the sensation seeking type of person. Also, while I have never visited the UK I have been told they sell cigarettes out of vending machines. Dont fact check me on that!

Training Set
```{r}
trainpredrf = predict(final_rf_fit, train)
head(trainpredrf)
```


```{r}
confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")
```
Testing Set
```{r}
testpredrf = predict(final_rf_fit, test)
head(testpredrf)
```

```{r}
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```
The model performs well on the training set, but when tested on the testing set it drops by 13%. This does not make me feel great, as it is a considerable drop in accuracy. The data set could be unbalanced. 

### Task 7
I believe this model could be used to back anti-smoking commercials. I think this model would be especially useful in targeting specific markets like specific age range sensation seekers in the UK. A concern I would have is that the testing and training set represented different accuracys. Although there was a 13% drop from training to testing. I still think that a 71% accuracy is decent. 